---
title: "Beta Calibration"
author: "Jianhui Gao"
date: "2025-11-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warnings = FALSE)
## ---- Setup (safe defaults if not defined) ----
if (!exists("rho")) rho <- 0.4
if (!exists("n_labeled")) n_labeled <- 1e4
if (!exists("N_unlabeled")) N_unlabeled <- 0
if (!exists("prot_att_prevalence")) prot_att_prevalence <- 0.4

library(MASS)
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(CalibrationCurves)
library(betacal)
```

# Motivation

## Base Case

### Latent Variable Generation

Generate latent variable $Z$ as 

$$
Z_a = \beta_a^T\boldsymbol{X} + \eta_a^T\boldsymbol{W} + \text{interactions} + \text{nonlinear terms}
$$

```{r}
## ---- Your provided block (kept as-is) ----
# Total sample size.
N_total <- 1e4
# Dimension
p <- 16
# Generate Covariates
id_matrix <- diag(p)
ar_one_matrix <- rho^abs(row(id_matrix) - col(id_matrix))
sigma <- 3 * ar_one_matrix
mu <- rep(0, p)
covariates <- MASS::mvrnorm(N_total, mu, Sigma = sigma)
colnames(covariates) <- c(paste0("X_", 1:10), paste0("W_", 1:5), "A")
# Placeholder for Y
Y <- rep(NA, N_total)
A <- ifelse(covariates[, "A"] > qnorm(1 - prot_att_prevalence), 1, 0)
W <- covariates[, 11:15, drop = FALSE]
X <- covariates[,  1:10, drop = FALSE]
## ---- Step 2: build latent t and standardized z (as in your DGP) ----
t <- -0.5 +
  1.00 * X[, 1] + 0.80 * X[, 2] + 0.60 * X[, 3] + 0.50 * X[, 4] + 0.30 * X[, 5] +
  0.30 * W[, 1] + 0.20 * W[, 2] - 0.20 * W[, 3] +
  0.10 * X[, 1] * W[, 1] - 0.10 * X[, 5] * X[, 2] +
  0.12 * (X[, 3]^2) - 0.08 * (X[, 3]^3) / 3 + 0.10 * (W[, 1]^2)
z <- as.numeric(scale(t))
```
```{r, fig.width=6, fig.height=3, fig.align='center'}
# Plot Z

# First, it's easiest to put 't' into a data.frame for ggplot
plot_data <- data.frame(z = z, A = factor(A))

# Now, create the histogram
ggplot(plot_data, aes(x = z, fill = A)) +
  geom_histogram(position = "identity", alpha = 0.5, bins = 50) +
  labs(title = "Histogram of z by Protected Attribute A",
       x = "Latent Variable z",
       y = "Count",
       fill = "Protected Attribute (A)") +
  theme_minimal() +
  theme(legend.position = "bottom")


# Link functions
p_true <- numeric(N_total)
for (a in 0:1) {
  idx <- which(A == a)
  za <- z[idx]
  f <- function(d) mean(1 - exp(-exp(za + d))) - 0.30
  delta <- tryCatch(
    uniroot(f, interval = c(-20, 20))$root,
    error = function(e) uniroot(f, interval = c(-40, 40))$root
  )
  if (a == 0) {
    p_true[idx] <- 1 - exp(-exp(za + delta))
  } else {
    p_true[idx] <- pnorm(za + delta)
  }
}
p_true <- pmin(pmax(p_true, 1e-6), 1 - 1e-6)
```

### Inverse cloglog link

$$
P(Y = 1 | Z_a) = 1 - \exp(-\exp(Z_a))
$$


```{r, fig.width=4, fig.height=3, fig.align='center'}
# Plot p_true vs z by A
plot_data_p <- data.frame(z = z, p_true = p_true, A = factor(A))
ggplot(plot_data_p, aes(x = z, y = p_true, color = A)) +
  geom_point(alpha = 0.5) +
  labs(title = "True Probability by Protected Attributes",
       x = "Latent Variable z",
       y = "True Probability",
       color = "Protected Attribute (A)") +
  theme_minimal() +
  theme(legend.position = "bottom")

Y <- rbinom(N_total, 1, p_true)
```

### Model Fitting and Calibration

Regress Y ~ X only within each group A = 0 and A = 1 separately. Then calibrate the predicted probabilities using Beta Calibration within each group.

```{r}
# model fitting
indep <- data.frame(Y = Y, A = A, X)
model_0 <- glm(
  Y ~ ., family = binomial(),
  data = indep %>% filter(A == 0)
)

model_1 <- glm(
  Y ~ ., family = binomial(),
  data = indep %>% filter(A == 1)
)

# Generate test data
S <- NA_real_
S[A == 0] <- model_0$fitted.values
S[A == 1] <- model_1$fitted.values

# Evaluate Calibration

#val.prob(S[A==0], Y[A==0], logistic.cal = FALSE)
#val.prob(S[A==1], Y[A==1], logistic.cal = FALSE)

# Beta Calibration
m <- beta_calibration(S[A==0], Y[A==0])
p1 <- val.prob(beta_predict(S[A==0], m), Y[A==0], logistic.cal = FALSE)
m <- beta_calibration(S[A==1], Y[A==1])
p2 <- val.prob(beta_predict(S[A==1], m), Y[A==1], logistic.cal = FALSE)
ggpubr::ggarrange(p1, p2, ncol = 2, nrow = 1)
```

## Modified Case:
### Add another layer of nonlinearity to $g_a(z)$ to create mirror-opposite groups

Define 
$$
g_0(z) = 3.2*z + 7.5*exp(-(z - 0.3)^2 / (2 * 0.18^2))
$$
and
$$
g_1(z) = 3.2*z + 7.5*exp(-(z + 0.3)^2 / (2 * 0.18^2))
$$

The shift in groups is needed, otherwise the bias will be in the same direction for both groups, and canceled out for delta. Then use same link functions as before to get p_true.

```{r, fig.width=4, fig.height=3, fig.align='center'}
## ---- Coefficients + “gap” knob (A=1 is scaled-up mirror of A=0) ----
s0  <- 0.18
gap <- 0.8  # increase for more separation; A=1 gets (1+gap) scaling

## Base (single) coefficients: define the *shape* once
b_base        <- 3.20
amp_up_1_base <- 7.50
# If you later want multiple bumps/ripples, add them inside f() (kept off by default)
amp_down_base <- -7.50
amp_up_2_base <- 9.00
ripple_amp_base <- 0.60
ripple_k <- 4

## ---- Symmetry-preserving transform (shared for both groups) ----
delta  <- 0.10   # center shift (common)
kappa  <- 0.10   # cubic warp (common)
width_mult <- 1.00
s0a <- s0 * width_mult

## ---- Define the base shape f(zt) once; group 1 will be -f(-zt) * (1+gap) ----
f <- function(zt) {
  b_base * zt +
    amp_up_1_base * exp(-(zt - 0.30)^2 / (2 * s0a^2))
  # + amp_down_base * exp(-(zt - 0.00)^2 / (2 * s0a^2)) +
  #   amp_up_2_base * exp(-(zt - 0.30)^2 / (2 * s0a^2)) +
  #   ripple_amp_base * exp(-(zt / 0.50)^2) * sin(2 * pi * ripple_k * plogis(zt))
}
f2 <- function(zt) {
  b_base * zt +
    amp_up_1_base * exp(-(zt + 0.30)^2 / (2 * s0a^2))
  # + amp_down_base * exp(-(zt - 0.00)^2 / (2 * s0a^2)) +
  #   amp_up_2_base * exp(-(zt - 0.30)^2 / (2 * s0a^2)) +
  #   ripple_amp_base * exp(-(zt / 0.50)^2) * sin(2 * pi * ripple_k * plogis(zt))
}

## ---- Nonlinear index g(z, A) with enforced anti-symmetry & scaling ----
g <- numeric(N_total)
zt_all <- z 
#+ kappa * (z^3) + delta  # same transform for everyone

idx0 <- which(A == 0)
idx1 <- which(A == 1)

g[idx0] <- f(zt_all[idx0])                    # A = 0
g[idx1] <- f2(zt_all[idx1])      # A = 1 is mirrored & scaled

## ---- Plot g vs z by A ----
plot_data_g <- data.frame(z = z, g = g, A = factor(A))
ggplot(plot_data_g, aes(x = z, y = g, color = A)) +
  geom_point(alpha = 0.5) +
  labs(title = "",
       x = "Standardized Latent Variable z",
       y = "Nonlinear Index g",
       color = "Protected Attribute (A)") +
  theme_minimal() +
  theme(legend.position = "bottom")
```
```{r, fig.width=4, fig.height=3, fig.align='center'}
p_true <- numeric(N_total)
for (a in 0:1) {
  idx <- which(A == a)
  ga <- g[idx]
  f <- function(d) mean(1 - exp(-exp(ga + d))) - 0.30
  delta <- tryCatch(
    uniroot(f, interval = c(-20, 20))$root,
    error = function(e) uniroot(f, interval = c(-40, 40))$root
  )
  if (a == 0) {
    p_true[idx] <- 1 - exp(-exp(ga + delta))
  } else {
    p_true[idx] <- pnorm(ga + delta)
  }
}
p_true <- pmin(pmax(p_true, 1e-6), 1 - 1e-6)

# Plot p_true vs g by A
plot_data_p_g <- data.frame(g = g, p_true = p_true, A = factor(A))
ggplot(plot_data_p_g, aes(x = g, y = p_true, color = A)) +
  geom_point(alpha = 0.5) +
  labs(title = "True Probability by Protected Attributes",
       x = "Nonlinear Index g",
       y = "True Probability p_true",
       color = "Protected Attribute (A)") +
  theme_minimal() +
  theme(legend.position = "bottom")

Y <- rbinom(N_total, 1, p_true)
```

### Model Fitting and Calibration
Regress Y ~ X only within each group A = 0 and A = 1 separately. Then calibrate the predicted probabilities using Beta Calibration within each group.

```{r}
# model fitting
indep <- data.frame(Y = Y, A = A, X)
model_0 <- glm(
  Y ~ ., family = binomial(),
  data = indep %>% filter(A == 0)
)

model_1 <- glm(
  Y ~ ., family = binomial(),
  data = indep %>% filter(A == 1)
)

# Generate test data
S <- NA_real_
S[A == 0] <- model_0$fitted.values
S[A == 1] <- model_1$fitted.values

# Evaluate Calibration

#val.prob(S[A==0], Y[A==0], logistic.cal = FALSE)
#val.prob(S[A==1], Y[A==1], logistic.cal = FALSE)

# Beta Calibration
m <- beta_calibration(S[A==0], Y[A==0])
val.prob(beta_predict(S[A==0], m), Y[A==0], logistic.cal = FALSE)
m <- beta_calibration(S[A==1], Y[A==1])
val.prob(beta_predict(S[A==1], m), Y[A==1], logistic.cal = FALSE)
```

# Results

```{r}
result <- readRDS("Data/scenario 5_n=400.rds")
#result2 <- readRDS("Data/scenario 3_n=1000.rds")
oracle <- readRDS("Data/scenario 5_oracle.rds")

oracle_est <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  )
all_lengths <- sapply(result, length)

# 2. Find the indices where the length is not 5
problem_indices <- which(all_lengths != 5)

# 3. Remove those problematic entries from the result list
if (length(problem_indices) > 0) {
  result <- result[-problem_indices]
}

# Loop over methods =================================
methods <- names(result[[1]])

results_list <- list()

# Loop over each method to compute bias and MSE
for (method in methods) {
  # Create the estimation part (mean)
  est <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  # Create the variance part (variance)
  avar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$var)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "aVar"
    )
  
  evar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) var(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "eVar"
    )
  
  # Combine estimation and variance
  combined <- est %>%
    left_join(avar, by = c("Metric", "Group")) %>%
    left_join(evar, by = c("Metric", "Group")) %>%
    mutate(Method = method)
  
  # Calculate the bias and MSE by comparing with oracle
  combined <- combined %>%
    left_join(oracle_est %>% dplyr::select(Metric, Group, Est) %>% rename(Oracle_Est = Est), 
              by = c("Metric", "Group")) %>%
    mutate(bias = Est - Oracle_Est, MSE = bias^2 + eVar)
  
  # Append to the results list
  results_list[[method]] <- combined
}

# Combine all the results into one data frame
final_results <- bind_rows(results_list)

# Define the methods for comparison, excluding 'oracle' and 'sup'
methods <- methods[!methods %in% c("oracle", "sup")]

labels <- c("Infairness(S only)","Infairness(S and W)", "Ji et al. (2020)", "Interaction")
#labels <- methods

# Filter the MSE from sup_est as the baseline
sup_mse <- final_results %>%
  filter(Method == "sup") %>%
  dplyr::select(Metric, Group, MSE) %>%
  rename(sup_MSE = MSE)

# Initialize an empty data frame to store RE results
re_final <- data.frame()

# Loop over each method to compute RE relative to sup_est
for (i in seq_along(methods)) {
  
  # Get the current method and label
  method <- methods[i]
  label <- labels[i]
  
  # Get MSE of the current method
  method_mse <- final_results %>%
    filter(Method == method) %>%
    dplyr::select(Metric, Group, MSE)
  
  # Compute RE (sup_MSE / method_MSE)
  re_temp <- sup_mse %>%
    left_join(method_mse, by = c("Metric", "Group")) %>%
    mutate(RE = sup_MSE / MSE, 
           Method = label)  # Apply the label
  
  # Append to the re_final data frame
  re_final <- rbind(re_final, re_temp)
}



plot_data <- final_results %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  mutate(
    CI_lower = Est - 1.96 * sqrt(eVar),
    CI_upper = Est + 1.96 * sqrt(eVar),
    Metric_num = as.numeric(factor(Metric))  # numeric x for drawing segments
  ) %>%
  filter(Method != "oracle") %>%
  mutate(Method = case_when(
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction",
    Method == "sup"        ~ "Supervised"
  )) %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                            "Infairness(S and W)",
                                            "Interaction",
                                            "Ji et al. (2020)",
                                            "Supervised")))

oracle_df <- plot_data %>%
  distinct(Metric, Oracle_Est, Metric_num, Group) %>%
  mutate(
    x_start = Metric_num - 0.45,
    x_end   = Metric_num + 0.45
  )

ggplot(plot_data, aes(x = factor(Metric_num), y = Est, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Group) +
  geom_segment(data = oracle_df,
               aes(x = x_start, xend = x_end,
                   y = Oracle_Est, yend = Oracle_Est),
               inherit.aes = FALSE,
               color = "black") +
  scale_x_discrete(
    labels = parse(text = oracle_df[oracle_df$Group == "Delta",]$Metric)) +
  ylab("Estimated Value") +
  xlab("") +
  ggsci::scale_color_nejm() +
  theme_bw() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                position = position_dodge2(width = 1)) +
  theme(legend.position = "bottom")


re_final %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  filter(!Method == "Ji et al. (2020)") %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                            "Infairness(S and W)",
                                            "Interaction"))) %>%
  ggplot(aes(x = Metric, y = RE, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge2()) +
  theme_bw() +
  facet_wrap(~Group) +
  geom_abline(intercept = 1, slope = 0, linetype = "dashed", col = "red") +
  ylab("Relative Efficiency") +
  xlab("") +
  scale_x_discrete(labels = scales::parse_format()) +
  ggsci::scale_fill_nejm() +
  theme(legend.position = "bottom")


oracle_df <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  ) %>%
  data.frame()

coverage_result <- c()
methods <- c("sup", "ss_s", "ss_sw", "ss_Beta", "ss_sw_int")

for (method in methods) {
  
  est <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1]))ll[[method]]$est)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  var <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$var)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Var"
    )
  
  for(m in unique(est$Metric)){
    for(g in unique(est$Group)){
      est_temp <- est %>% filter(Metric == m, Group == g) %>% pull(Est)
      var_temp <- var %>% filter(Metric == m, Group == g) %>% pull(Var)
      
      coverage <- sum((est_temp - qnorm(0.975) * sqrt(var_temp) <= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"])) &
                        (est_temp + qnorm(0.975) * sqrt(var_temp) >= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"]))) / length(est_temp)
      
      coverage_result <- rbind(coverage_result, data.frame(Method = method, Metric = m, Group = g, Coverage = coverage))
    }
  }
  
}

coverage_result %>%
  mutate(Method = case_when(
    Method == "sup"        ~ "Supervised",
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction"
  )) %>%
  mutate(Method = factor(Method, levels = c(
    "Infairness(S only)",
    "Infairness(S and W)",
    "Interaction",
    "Ji et al. (2020)",
    "Supervised"))) %>%
  ggplot(aes(x = Group, y = Coverage, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Metric) +
  ylab("Coverage Probability") +
  xlab("") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ggsci::scale_color_nejm()

model_eval_list <- list()

for (method in methods) {
  eval <- do.call(rbind, lapply(result, function(ll) {
    if (!is.null(ll[[method]]) && !is.na(ll[[method]][1])) ll[[method]]$imp_quality else NULL
  })) %>%
    as.data.frame()
  
  if (nrow(eval) == 0L) next
  
  # --- Exclude Inf values ---
  eval[!is.finite(as.matrix(eval))] <- NA   # replaces Inf / -Inf with NA
  
  odd_index  <- seq(1, nrow(eval), by = 2)
  even_index <- seq(2, nrow(eval), by = 2)
  
  g0 <- colMeans(eval[odd_index, , drop = FALSE], na.rm = TRUE)
  g1 <- colMeans(eval[even_index, , drop = FALSE], na.rm = TRUE)
  
  tmp <- rbind(g0, g1) %>% as.data.frame()
  tmp$Group  <- rownames(tmp)
  tmp$Method <- method
  rownames(tmp) <- NULL
  tmp <- tmp[, c("Method", "Group", setdiff(names(tmp), c("Method", "Group")))]
  
  model_eval_list[[length(model_eval_list) + 1L]] <- tmp
}

model_eval <- dplyr::bind_rows(model_eval_list)
model_eval

```

