---
title: "Simulation Revisit"
author: "Jianhui Gao"
date: "2025-10-26"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r, setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(SSFairnessAudit)
source("DataGeneration.R")
prev <- 0.4
rho <- 0.4
threshold <- 0.5
knitr::opts_chunk$set(echo = FALSE, fig.width = 12, fig.height = 4)
```

```{=tex}
\tableofcontents

```

\newpage

# Data Generation

To generate $(X^T , W^T , A)^T$ , we first simulated 16 random variables
from a zero-mean multivariate normal distribution with a first-order
autoregressive covariance structure. The covariance between the k-th and
l-th random variables was set to $3(0.4)^{|k-l|}$. The first 10 random
variables were used as the features, $X$, for model training, and the
next 5 as auxiliary covariates, $W$, for validation. The final random
variable was used to obtain the protected attribute, $A$, and set to 1
if the normally generated value exceeded 0.253 and set to 0 otherwise.
The outcome $Y$ was generated under two diﬀerent scenarios to induce
varying degrees of misspecification into the ML and imputation models: 

- Scenario 1 (Old): Logistic activation function with non-linear inputs
$$P(Y = 1 | X, W, A= a) = expit\{-2.3 + \beta_a^TX + \gamma_a^TW + 0.2X_2^2- 0.1X_3^2 + 0.1X_5X_6\}$$


- Scenario 2 (Old): Non-logistic activation function with non-linear hidden
layer $$P (Y = 1 | X, W, A= a) = exp\{-(1.3 + \beta_a^TX + \gamma_a^T
W)^2\}$$

- Scenario 3 (New): Tree Structure

$\eta = \beta^TX + \gamma^T + X_4^2 - X_5^2 + 0.3X_1W_1 - 0.3X_2W_2 + 0.2X_3W_3 + \tanh(0.4W_5) - \tanh(0.2X_4) + 0.1\sin(2W_4X_5)$ and $\psi = \beta^TX + \gamma^T - 0.4W_4^2 + 0.3W_5^2 + 0.2X_1W_1 - 0.2X_2W_2 + 0.1X_3W_3 + \tanh(0.3X_5) - \tanh(0.1W_4) + 0.1\cos(2X_4W_5)$. Then generate S as:

$$
S = \begin{cases}
0.8 & if A = 0, \eta_a > 0.5, \psi_a > 0.3 \\
0.6 & if A = 0, \eta_a > 0.5, \psi_a \leq 0.3 \\
0.4 & if A = 0, \eta_a \leq 0.5, \psi_a > 0.3 \\
0.2 & if A = 0, \eta_a \leq 0.5, \psi_a \leq 0.3 \\
0.7 & if A = 1, \eta_a > 0.3, \psi_a > 0.5 \\
0.5 & if A = 1, \eta_a > 0.3, \psi_a \leq 0.5 \\
0.3 & if A = 1, \eta_a \leq 0.3, \psi_a > 0.5 \\
0.1 & if A = 1, \eta_a \leq 0.3, \psi_a \leq 0.5 \\
\end{cases}
$$

Finally $P(Y = 1 \mid X, W, A = a) = S$




- Scenario 4 (New): Log-log activation function with non-monotone effect

$$P(Y = 1 | X, W, A= a) = 1 - exp\{-exp(g_a(Z))\}$$

where 
$$
Z = -0.5 + \beta_a^TX + \gamma_a^TW + 0.1X_1W_1 - 0.1X_2X_5 + 0.12X_3^2 - 0.08X_3^3/3 + 0.10W_1^2
$$
and

$$
g_a(z)\;=\; b_a\,z
\;+\; u_{1a}\,\exp\!\Big\{-\tfrac{(z+0.30)^2}{2\,s_0^2}\Big\}
\;+\; d_a\,\exp\!\Big\{-\tfrac{z^2}{2\,s_0^2}\Big\}
\;+\; u_{2a}\,\exp\!\Big\{-\tfrac{(z-0.30)^2}{2\,s_0^2}\Big\}
\;+\; r_a\,\exp\!\big\{-(z/0.50)^2\big\}\,\sin\!\big(2\pi k\,\sigma(z)\big)
$$

# Scenario 1 

```{r}
result <- readRDS("Data/scenario 1_n=400.rds")
oracle <- readRDS("Data/scenario 1_oracle.rds")

oracle_est <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  )

all_lengths <- sapply(result, length)

# 2. Find the indices where the length is not 5
problem_indices <- which(all_lengths != 5)

# 3. Remove those problematic entries from the result list
if (length(problem_indices) > 0) {
  result <- result[-problem_indices]
}
```

```{r}
# Loop over methods =================================
methods <- names(result[[1]])

results_list <- list()

# Loop over each method to compute bias and MSE
for (method in methods) {
  
  # Create the estimation part (mean)
  est <- do.call(rbind, lapply(result, function(ll) if(!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  # Create the variance part (variance)
  avar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$var)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "aVar"
    )
  
  evar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) var(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "eVar"
    )
  
  # Combine estimation and variance
  combined <- est %>%
    left_join(avar, by = c("Metric", "Group")) %>%
    left_join(evar, by = c("Metric", "Group")) %>%
    mutate(Method = method)
  
  # Calculate the bias and MSE by comparing with oracle
  combined <- combined %>%
    left_join(oracle_est %>% select(Metric, Group, Est) %>% rename(Oracle_Est = Est), 
              by = c("Metric", "Group")) %>%
    mutate(bias = Est - Oracle_Est, MSE = bias^2 + eVar)
  
  # Append to the results list
  results_list[[method]] <- combined
}

# Combine all the results into one data frame
final_results <- bind_rows(results_list)

# Define the methods for comparison, excluding 'oracle' and 'sup'
methods <- methods[!methods %in% c("oracle", "sup")]

labels <- c("Infairness(S only)","Infairness(S and W)", "Ji et al. (2020)", "Interaction")
#labels <- methods

# Filter the MSE from sup_est as the baseline
sup_mse <- final_results %>%
  filter(Method == "sup") %>%
  select(Metric, Group, MSE) %>%
  rename(sup_MSE = MSE)

# Initialize an empty data frame to store RE results
re_final <- data.frame()

# Loop over each method to compute RE relative to sup_est
for (i in seq_along(methods)) {
  
  # Get the current method and label
  method <- methods[i]
  label <- labels[i]
  
  # Get MSE of the current method
  method_mse <- final_results %>%
    filter(Method == method) %>%
    select(Metric, Group, MSE)
  
  # Compute RE (sup_MSE / method_MSE)
  re_temp <- sup_mse %>%
    left_join(method_mse, by = c("Metric", "Group")) %>%
    mutate(RE = sup_MSE / MSE, 
           Method = label)  # Apply the label
  
  # Append to the re_final data frame
  re_final <- rbind(re_final, re_temp)
}
```

```{r}
plot_data <- final_results %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  mutate(
    CI_lower = Est - 1.96 * sqrt(eVar),
    CI_upper = Est + 1.96 * sqrt(eVar),
    Metric_num = as.numeric(factor(Metric))  # numeric x for drawing segments
  ) %>%
  filter(Method != "oracle") %>%
  mutate(Method = case_when(
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction",
    Method == "sup"        ~ "Supervised"
  )) %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised")))

oracle_df <- plot_data %>%
  distinct(Metric, Oracle_Est, Metric_num, Group) %>%
  mutate(
    x_start = Metric_num - 0.45,
    x_end   = Metric_num + 0.45
  )

ggplot(plot_data, aes(x = factor(Metric_num), y = Est, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Group) +
  geom_segment(data = oracle_df,
               aes(x = x_start, xend = x_end,
                   y = Oracle_Est, yend = Oracle_Est),
               inherit.aes = FALSE,
               color = "black") +
  scale_x_discrete(
    labels = parse(text = oracle_df[oracle_df$Group == "Delta",]$Metric)) +
  ylab("Estimated Value") +
  xlab("") +
  ggsci::scale_color_nejm() +
  theme_bw() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                position = position_dodge2(width = 1)) +
  theme(legend.position = "bottom")
```

```{r}
re_final %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  filter(!Method == "Ji et al. (2020)") %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction"))) %>%
  ggplot(aes(x = Metric, y = RE, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge2()) +
  theme_bw() +
  facet_wrap(~Group) +
  geom_abline(intercept = 1, slope = 0, linetype = "dashed", col = "red") +
  ylab("Relative Efficiency") +
  xlab("") +
  scale_x_discrete(labels = scales::parse_format()) +
  ggsci::scale_fill_nejm() +
  theme(legend.position = "bottom")
```



```{r, fig.width = 12, fig.height = 6}
oracle_df <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  ) %>%
  data.frame()

coverage_result <- c()
methods <- c("sup", "ss_s", "ss_sw", "ss_Beta", "ss_sw_int")

for (method in methods) {
  
  est <- do.call(rbind, lapply(result, function(ll) ll[[method]]$est)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  var <- do.call(rbind, lapply(result, function(ll) ll[[method]]$var)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Var"
    )
  
  for(m in unique(est$Metric)){
    for(g in unique(est$Group)){
      est_temp <- est %>% filter(Metric == m, Group == g) %>% pull(Est)
      var_temp <- var %>% filter(Metric == m, Group == g) %>% pull(Var)
      
      coverage <- sum((est_temp - qnorm(0.975) * sqrt(var_temp) <= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"])) &
                        (est_temp + qnorm(0.975) * sqrt(var_temp) >= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"]))) / length(est_temp)
      
      coverage_result <- rbind(coverage_result, data.frame(Method = method, Metric = m, Group = g, Coverage = coverage))
    }
  }
  
}

coverage_result %>%
  mutate(Method = case_when(
    Method == "sup"        ~ "Supervised",
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction"
  )) %>%
  mutate(Method = factor(Method, levels = c(
                                      "Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised"))) %>%
  ggplot(aes(x = Group, y = Coverage, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Metric) +
  ylab("Coverage Probability") +
  xlab("") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ggsci::scale_color_nejm()
```

```{r}

model_eval_list <- list()

for (method in methods) {
  eval <- do.call(rbind, lapply(result, function(ll) {
    if (!is.null(ll[[method]]) && !is.na(ll[[method]][1])) ll[[method]]$imp_quality else NULL
  })) %>%
    as.data.frame()
  
  if (nrow(eval) == 0L) next
  
  # --- Exclude Inf values ---
  eval[!is.finite(as.matrix(eval))] <- NA   # replaces Inf / -Inf with NA
  
  odd_index  <- seq(1, nrow(eval), by = 2)
  even_index <- seq(2, nrow(eval), by = 2)
  
  g0 <- colMeans(eval[odd_index, , drop = FALSE], na.rm = TRUE)
  g1 <- colMeans(eval[even_index, , drop = FALSE], na.rm = TRUE)
  
  tmp <- rbind(g0, g1) %>% as.data.frame()
  tmp$Group  <- rownames(tmp)
  tmp$Method <- method
  rownames(tmp) <- NULL
  tmp <- tmp[, c("Method", "Group", setdiff(names(tmp), c("Method", "Group")))]
  
  model_eval_list[[length(model_eval_list) + 1L]] <- tmp
}

model_eval <- dplyr::bind_rows(model_eval_list)
model_eval
```



# Scenario 2

```{r}
result <- readRDS("Data/scenario 2_n=400.rds")
oracle <- readRDS("Data/scenario 2_oracle.rds")

oracle_est <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  )


```

```{r}
# Loop over methods =================================
methods <- names(result[[1]])

results_list <- list()

# Loop over each method to compute bias and MSE
for (method in methods) {
  
  # Create the estimation part (mean)
  est <- do.call(rbind, lapply(result, function(ll) if(!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  # Create the variance part (variance)
  avar <- do.call(rbind, lapply(result, function(ll) if(!is.na(ll[[method]][1])) ll[[method]]$var)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "aVar"
    )
  
  evar <- do.call(rbind, lapply(result, function(ll) if(!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) var(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "eVar"
    )
  
  # Combine estimation and variance
  combined <- est %>%
    left_join(avar, by = c("Metric", "Group")) %>%
    left_join(evar, by = c("Metric", "Group")) %>%
    mutate(Method = method)
  
  # Calculate the bias and MSE by comparing with oracle
  combined <- combined %>%
    left_join(oracle_est %>% select(Metric, Group, Est) %>% rename(Oracle_Est = Est), 
              by = c("Metric", "Group")) %>%
    mutate(bias = Est - Oracle_Est, MSE = bias^2 + eVar)
  
  # Append to the results list
  results_list[[method]] <- combined
}

# Combine all the results into one data frame
final_results <- bind_rows(results_list)

# Define the methods for comparison, excluding 'oracle' and 'sup'
methods <- methods[!methods %in% c("oracle", "sup")]

labels <- c("Infairness(S only)","Infairness(S and W)", "Ji et al. (2020)", "Interaction")
#labels <- methods

# Filter the MSE from sup_est as the baseline
sup_mse <- final_results %>%
  filter(Method == "sup") %>%
  select(Metric, Group, MSE) %>%
  rename(sup_MSE = MSE)

# Initialize an empty data frame to store RE results
re_final <- data.frame()

# Loop over each method to compute RE relative to sup_est
for (i in seq_along(methods)) {
  
  # Get the current method and label
  method <- methods[i]
  label <- labels[i]
  
  # Get MSE of the current method
  method_mse <- final_results %>%
    filter(Method == method) %>%
    select(Metric, Group, MSE)
  
  # Compute RE (sup_MSE / method_MSE)
  re_temp <- sup_mse %>%
    left_join(method_mse, by = c("Metric", "Group")) %>%
    mutate(RE = sup_MSE / MSE, 
           Method = label)  # Apply the label
  
  # Append to the re_final data frame
  re_final <- rbind(re_final, re_temp)
}
```

```{r}
plot_data <- final_results %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  mutate(
    CI_lower = Est - 1.96 * sqrt(eVar),
    CI_upper = Est + 1.96 * sqrt(eVar),
    Metric_num = as.numeric(factor(Metric))  # numeric x for drawing segments
  ) %>%
  filter(Method != "oracle") %>%
  mutate(Method = case_when(
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction",
    Method == "sup"        ~ "Supervised"
  )) %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised")))

oracle_df <- plot_data %>%
  distinct(Metric, Oracle_Est, Metric_num, Group) %>%
  mutate(
    x_start = Metric_num - 0.45,
    x_end   = Metric_num + 0.45
  )

ggplot(plot_data, aes(x = factor(Metric_num), y = Est, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Group) +
  geom_segment(data = oracle_df,
               aes(x = x_start, xend = x_end,
                   y = Oracle_Est, yend = Oracle_Est),
               inherit.aes = FALSE,
               color = "black") +
  scale_x_discrete(
    labels = parse(text = oracle_df[oracle_df$Group == "Delta",]$Metric)) +
  ylab("Estimated Value") +
  xlab("") +
  ggsci::scale_color_nejm() +
  theme_bw() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                position = position_dodge2(width = 1)) +
  theme(legend.position = "bottom")
```

```{r}
re_final %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  filter(!Method == "Ji et al. (2020)") %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction"))) %>%
  ggplot(aes(x = Metric, y = RE, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge2()) +
  theme_bw() +
  facet_wrap(~Group) +
  geom_abline(intercept = 1, slope = 0, linetype = "dashed", col = "red") +
  ylab("Relative Efficiency") +
  xlab("") +
  scale_x_discrete(labels = scales::parse_format()) +
  ggsci::scale_fill_nejm() +
  theme(legend.position = "bottom")
```

```{r, fig.width = 12, fig.height = 6}
oracle_df <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  ) %>%
  data.frame()

coverage_result <- c()
methods <- c("sup", "ss_s", "ss_sw", "ss_Beta", "ss_sw_int")

for (method in methods) {
  
  est <- do.call(rbind, lapply(result, function(ll) ll[[method]]$est)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  var <- do.call(rbind, lapply(result, function(ll) ll[[method]]$var)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Var"
    )
  
  for(m in unique(est$Metric)){
    for(g in unique(est$Group)){
      est_temp <- est %>% filter(Metric == m, Group == g) %>% pull(Est)
      var_temp <- var %>% filter(Metric == m, Group == g) %>% pull(Var)
      
      coverage <- sum((est_temp - qnorm(0.975) * sqrt(var_temp) <= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"])) &
                        (est_temp + qnorm(0.975) * sqrt(var_temp) >= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"]))) / length(est_temp)
      
      coverage_result <- rbind(coverage_result, data.frame(Method = method, Metric = m, Group = g, Coverage = coverage))
    }
  }
  
}

coverage_result %>%
  mutate(Method = case_when(
    Method == "sup"        ~ "Supervised",
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction"
  )) %>%
  mutate(Method = factor(Method, levels = c(
                                      "Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised"))) %>%
  ggplot(aes(x = Group, y = Coverage, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Metric) +
  ylab("Coverage Probability") +
  xlab("") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ggsci::scale_color_nejm()
```

```{r}

model_eval_list <- list()

for (method in methods) {
  eval <- do.call(rbind, lapply(result, function(ll) {
    if (!is.null(ll[[method]]) && !is.na(ll[[method]][1])) ll[[method]]$imp_quality else NULL
  })) %>%
    as.data.frame()
  
  if (nrow(eval) == 0L) next
  
  # --- Exclude Inf values ---
  eval[!is.finite(as.matrix(eval))] <- NA   # replaces Inf / -Inf with NA
  
  odd_index  <- seq(1, nrow(eval), by = 2)
  even_index <- seq(2, nrow(eval), by = 2)
  
  g0 <- colMeans(eval[odd_index, , drop = FALSE], na.rm = TRUE)
  g1 <- colMeans(eval[even_index, , drop = FALSE], na.rm = TRUE)
  
  tmp <- rbind(g0, g1) %>% as.data.frame()
  tmp$Group  <- rownames(tmp)
  tmp$Method <- method
  rownames(tmp) <- NULL
  tmp <- tmp[, c("Method", "Group", setdiff(names(tmp), c("Method", "Group")))]
  
  model_eval_list[[length(model_eval_list) + 1L]] <- tmp
}

model_eval <- dplyr::bind_rows(model_eval_list)
model_eval
```



# Scenario 3 (Tree Model)

```{r}
result <- readRDS("Data/tree_n=400.rds")
oracle <- readRDS("Data/tree_oracle.rds")

oracle_est <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  )
all_lengths <- sapply(result, length)

# 2. Find the indices where the length is not 5
problem_indices <- which(all_lengths != 5)

# 3. Remove those problematic entries from the result list
if (length(problem_indices) > 0) {
  result <- result[-problem_indices]
}
```

```{r}
# Loop over methods =================================
methods <- names(result[[1]])

results_list <- list()

# Loop over each method to compute bias and MSE
for (method in methods) {
  
  # Create the estimation part (mean)
  est <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1]))  ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  # Create the variance part (variance)
  avar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1]))  ll[[method]]$var)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "aVar"
    )
  
  evar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) var(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "eVar"
    )
  
  # Combine estimation and variance
  combined <- est %>%
    left_join(avar, by = c("Metric", "Group")) %>%
    left_join(evar, by = c("Metric", "Group")) %>%
    mutate(Method = method)
  
  # Calculate the bias and MSE by comparing with oracle
  combined <- combined %>%
    left_join(oracle_est %>% select(Metric, Group, Est) %>% rename(Oracle_Est = Est), 
              by = c("Metric", "Group")) %>%
    mutate(bias = Est - Oracle_Est, MSE = bias^2 + eVar)
  
  # Append to the results list
  results_list[[method]] <- combined
}

# Combine all the results into one data frame
final_results <- bind_rows(results_list)

# Define the methods for comparison, excluding 'oracle' and 'sup'
methods <- methods[!methods %in% c("oracle", "sup")]

labels <- c("Infairness(S only)","Infairness(S and W)", "Ji et al. (2020)", "Interaction")
#labels <- methods

# Filter the MSE from sup_est as the baseline
sup_mse <- final_results %>%
  filter(Method == "sup") %>%
  select(Metric, Group, MSE) %>%
  rename(sup_MSE = MSE)

# Initialize an empty data frame to store RE results
re_final <- data.frame()

# Loop over each method to compute RE relative to sup_est
for (i in seq_along(methods)) {
  
  # Get the current method and label
  method <- methods[i]
  label <- labels[i]
  
  # Get MSE of the current method
  method_mse <- final_results %>%
    filter(Method == method) %>%
    select(Metric, Group, MSE)
  
  # Compute RE (sup_MSE / method_MSE)
  re_temp <- sup_mse %>%
    left_join(method_mse, by = c("Metric", "Group")) %>%
    mutate(RE = sup_MSE / MSE, 
           Method = label)  # Apply the label
  
  # Append to the re_final data frame
  re_final <- rbind(re_final, re_temp)
}
```

```{r}
plot_data <- final_results %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  mutate(
    CI_lower = Est - 1.96 * sqrt(eVar),
    CI_upper = Est + 1.96 * sqrt(eVar),
    Metric_num = as.numeric(factor(Metric))  # numeric x for drawing segments
  ) %>%
  filter(Method != "oracle") %>%
  mutate(Method = case_when(
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction",
    Method == "sup"        ~ "Supervised"
  )) %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised")))

oracle_df <- plot_data %>%
  distinct(Metric, Oracle_Est, Metric_num, Group) %>%
  mutate(
    x_start = Metric_num - 0.45,
    x_end   = Metric_num + 0.45
  )

ggplot(plot_data, aes(x = factor(Metric_num), y = Est, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Group) +
  geom_segment(data = oracle_df,
               aes(x = x_start, xend = x_end,
                   y = Oracle_Est, yend = Oracle_Est),
               inherit.aes = FALSE,
               color = "black") +
  scale_x_discrete(
    labels = parse(text = oracle_df[oracle_df$Group == "Delta",]$Metric)) +
  ylab("Estimated Value") +
  xlab("") +
  ggsci::scale_color_nejm() +
  theme_bw() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                position = position_dodge2(width = 1)) +
  theme(legend.position = "bottom")
```

```{r}
re_final %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  filter(!Method == "Ji et al. (2020)") %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction"))) %>%
  ggplot(aes(x = Metric, y = RE, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge2()) +
  theme_bw() +
  facet_wrap(~Group) +
  geom_abline(intercept = 1, slope = 0, linetype = "dashed", col = "red") +
  ylab("Relative Efficiency") +
  xlab("") +
  scale_x_discrete(labels = scales::parse_format()) +
  ggsci::scale_fill_nejm() +
  theme(legend.position = "bottom")
```

```{r, fig.width = 12, fig.height = 6}
oracle_df <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  ) %>%
  data.frame()

coverage_result <- c()
methods <- c("sup", "ss_s", "ss_sw", "ss_Beta", "ss_sw_int")

for (method in methods) {
  
  est <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  var <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$var)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Var"
    )
  
  for(m in unique(est$Metric)){
    for(g in unique(est$Group)){
      est_temp <- est %>% filter(Metric == m, Group == g) %>% pull(Est)
      var_temp <- var %>% filter(Metric == m, Group == g) %>% pull(Var)
      
      coverage <- sum((est_temp - qnorm(0.975) * sqrt(var_temp) <= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"])) &
                        (est_temp + qnorm(0.975) * sqrt(var_temp) >= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"]))) / length(est_temp)
      
      coverage_result <- rbind(coverage_result, data.frame(Method = method, Metric = m, Group = g, Coverage = coverage))
    }
  }
  
}

coverage_result %>%
  mutate(Method = case_when(
    Method == "sup"        ~ "Supervised",
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction"
  )) %>%
  mutate(Method = factor(Method, levels = c(
                                      "Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised"))) %>%
  ggplot(aes(x = Group, y = Coverage, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Metric) +
  ylab("Coverage Probability") +
  xlab("") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ggsci::scale_color_nejm()
```

```{r}
# percentage bias
final_results %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  mutate(Percent_Bias = (bias / Oracle_Est) * 100) %>%
  ggplot(aes(x = Metric, y = Percent_Bias, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge2()) +
  theme_bw() +
  facet_wrap(~Group) +
  geom_hline(yintercept = 0, linetype = "dashed", col = "red") +
  ylab("Percentage Bias (%)") +
  xlab("") +
  scale_x_discrete(labels = scales::parse_format()) +
  ggsci::scale_fill_nejm() +
  theme(legend.position = "bottom")
```


```{r}

model_eval_list <- list()

for (method in methods) {
  eval <- do.call(rbind, lapply(result, function(ll) {
    if (!is.null(ll[[method]]) && !is.na(ll[[method]][1])) ll[[method]]$imp_quality else NULL
  })) %>%
    as.data.frame()
  
  if (nrow(eval) == 0L) next
  
  # --- Exclude Inf values ---
  eval[!is.finite(as.matrix(eval))] <- NA   # replaces Inf / -Inf with NA
  
  odd_index  <- seq(1, nrow(eval), by = 2)
  even_index <- seq(2, nrow(eval), by = 2)
  
  g0 <- colMeans(eval[odd_index, , drop = FALSE], na.rm = TRUE)
  g1 <- colMeans(eval[even_index, , drop = FALSE], na.rm = TRUE)
  
  tmp <- rbind(g0, g1) %>% as.data.frame()
  tmp$Group  <- rownames(tmp)
  tmp$Method <- method
  rownames(tmp) <- NULL
  tmp <- tmp[, c("Method", "Group", setdiff(names(tmp), c("Method", "Group")))]
  
  model_eval_list[[length(model_eval_list) + 1L]] <- tmp
}

model_eval <- dplyr::bind_rows(model_eval_list)
model_eval
```



# Scenario 4

```{r}
result <- readRDS("Data/scenario 5_n=400.rds")
oracle <- readRDS("Data/scenario 5_oracle.rds")

oracle_est <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  )
all_lengths <- sapply(result, length)

# 2. Find the indices where the length is not 5
problem_indices <- which(all_lengths != 5)

# 3. Remove those problematic entries from the result list
if (length(problem_indices) > 0) {
  result <- result[-problem_indices]
}
```

```{r}
# Loop over methods =================================
methods <- names(result[[1]])

results_list <- list()

# Loop over each method to compute bias and MSE
for (method in methods) {
  
  # Create the estimation part (mean)
  est <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1]))
                               ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  # Create the variance part (variance)
  avar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1]))ll[[method]]$var)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) mean(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "aVar"
    )
  
  evar <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1]))ll[[method]]$est)) %>%
    group_by(Metric) %>%
    summarise_all(function(x) var(x, na.rm = TRUE)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "eVar"
    )
  
  # Combine estimation and variance
  combined <- est %>%
    left_join(avar, by = c("Metric", "Group")) %>%
    left_join(evar, by = c("Metric", "Group")) %>%
    mutate(Method = method)
  
  # Calculate the bias and MSE by comparing with oracle
  combined <- combined %>%
    left_join(oracle_est %>% select(Metric, Group, Est) %>% rename(Oracle_Est = Est), 
              by = c("Metric", "Group")) %>%
    mutate(bias = Est - Oracle_Est, MSE = bias^2 + eVar)
  
  # Append to the results list
  results_list[[method]] <- combined
}

# Combine all the results into one data frame
final_results <- bind_rows(results_list)

# Define the methods for comparison, excluding 'oracle' and 'sup'
methods <- methods[!methods %in% c("oracle", "sup")]

labels <- c("Infairness(S only)","Infairness(S and W)", "Ji et al. (2020)", "Interaction")
#labels <- methods

# Filter the MSE from sup_est as the baseline
sup_mse <- final_results %>%
  filter(Method == "sup") %>%
  select(Metric, Group, MSE) %>%
  rename(sup_MSE = MSE)

# Initialize an empty data frame to store RE results
re_final <- data.frame()

# Loop over each method to compute RE relative to sup_est
for (i in seq_along(methods)) {
  
  # Get the current method and label
  method <- methods[i]
  label <- labels[i]
  
  # Get MSE of the current method
  method_mse <- final_results %>%
    filter(Method == method) %>%
    select(Metric, Group, MSE)
  
  # Compute RE (sup_MSE / method_MSE)
  re_temp <- sup_mse %>%
    left_join(method_mse, by = c("Metric", "Group")) %>%
    mutate(RE = sup_MSE / MSE, 
           Method = label)  # Apply the label
  
  # Append to the re_final data frame
  re_final <- rbind(re_final, re_temp)
}
```

```{r}
plot_data <- final_results %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  mutate(
    CI_lower = Est - 1.96 * sqrt(eVar),
    CI_upper = Est + 1.96 * sqrt(eVar),
    Metric_num = as.numeric(factor(Metric))  # numeric x for drawing segments
  ) %>%
  filter(Method != "oracle") %>%
  mutate(Method = case_when(
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction",
    Method == "sup"        ~ "Supervised"
  )) %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised")))

oracle_df <- plot_data %>%
  distinct(Metric, Oracle_Est, Metric_num, Group) %>%
  mutate(
    x_start = Metric_num - 0.45,
    x_end   = Metric_num + 0.45
  )

ggplot(plot_data, aes(x = factor(Metric_num), y = Est, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Group) +
  geom_segment(data = oracle_df,
               aes(x = x_start, xend = x_end,
                   y = Oracle_Est, yend = Oracle_Est),
               inherit.aes = FALSE,
               color = "black") +
  scale_x_discrete(
    labels = parse(text = oracle_df[oracle_df$Group == "Delta",]$Metric)) +
  ylab("Estimated Value") +
  xlab("") +
  ggsci::scale_color_nejm() +
  theme_bw() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                position = position_dodge2(width = 1)) +
  theme(legend.position = "bottom")
```

```{r}
re_final %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  filter(!Method == "Ji et al. (2020)") %>%
  mutate(Method = factor(Method, levels = c("Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction"))) %>%
  ggplot(aes(x = Metric, y = RE, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge2()) +
  theme_bw() +
  facet_wrap(~Group) +
  geom_abline(intercept = 1, slope = 0, linetype = "dashed", col = "red") +
  ylab("Relative Efficiency") +
  xlab("") +
  scale_x_discrete(labels = scales::parse_format()) +
  ggsci::scale_fill_nejm() +
  theme(legend.position = "bottom")
```

```{r, fig.width = 12, fig.height = 6}
oracle_df <- oracle$est %>%
  pivot_longer(
    cols = -c(Metric),
    names_to = "Group",
    values_to = "Est"
  ) %>%
  data.frame()

coverage_result <- c()
methods <- c("sup", "ss_s", "ss_sw", "ss_Beta", "ss_sw_int")

for (method in methods) {
  
  est <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$est)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Est"
    )
  
  var <- do.call(rbind, lapply(result, function(ll) if (!is.na(ll[[method]][1])) ll[[method]]$var)) %>%
    pivot_longer(
      cols = -c(Metric),
      names_to = "Group",
      values_to = "Var"
    )
  
  for(m in unique(est$Metric)){
    for(g in unique(est$Group)){
      est_temp <- est %>% filter(Metric == m, Group == g) %>% pull(Est)
      var_temp <- var %>% filter(Metric == m, Group == g) %>% pull(Var)
      
      coverage <- sum((est_temp - qnorm(0.975) * sqrt(var_temp) <= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"])) &
                        (est_temp + qnorm(0.975) * sqrt(var_temp) >= as.numeric(oracle_df [oracle_df $Metric == m & oracle_df $Group == g, "Est"]))) / length(est_temp)
      
      coverage_result <- rbind(coverage_result, data.frame(Method = method, Metric = m, Group = g, Coverage = coverage))
    }
  }
  
}

coverage_result %>%
  mutate(Method = case_when(
    Method == "sup"        ~ "Supervised",
    Method == "ss_s"       ~ "Infairness(S only)",
    Method == "ss_sw"      ~ "Infairness(S and W)",
    Method == "ss_Beta"    ~ "Ji et al. (2020)",
    Method == "ss_sw_int"  ~ "Interaction"
  )) %>%
  mutate(Method = factor(Method, levels = c(
                                      "Infairness(S only)",
                                      "Infairness(S and W)",
                                      "Interaction",
                                      "Ji et al. (2020)",
                                      "Supervised"))) %>%
  ggplot(aes(x = Group, y = Coverage, color = Method)) +
  geom_point(position = position_dodge2(width = 1), size = 2) +
  facet_wrap(~Metric) +
  ylab("Coverage Probability") +
  xlab("") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ggsci::scale_color_nejm()
```

```{r}
# percentage bias
final_results %>%
  filter(!Metric %in% c("TNR", "FNR")) %>%
  mutate(Percent_Bias = (bias / Oracle_Est) * 100) %>%
  ggplot(aes(x = Metric, y = Percent_Bias, fill = Method)) +
  geom_bar(stat = "identity", position = position_dodge2()) +
  theme_bw() +
  facet_wrap(~Group) +
  geom_hline(yintercept = 0, linetype = "dashed", col = "red") +
  ylab("Percentage Bias (%)") +
  xlab("") +
  scale_x_discrete(labels = scales::parse_format()) +
  ggsci::scale_fill_nejm() +
  theme(legend.position = "bottom")
```

```{r}

model_eval_list <- list()

for (method in methods) {
  eval <- do.call(rbind, lapply(result, function(ll) {
    if (!is.null(ll[[method]]) && !is.na(ll[[method]][1])) ll[[method]]$imp_quality else NULL
  })) %>%
    as.data.frame()
  
  if (nrow(eval) == 0L) next
  
  # --- Exclude Inf values ---
  eval[!is.finite(as.matrix(eval))] <- NA   # replaces Inf / -Inf with NA
  
  odd_index  <- seq(1, nrow(eval), by = 2)
  even_index <- seq(2, nrow(eval), by = 2)
  
  g0 <- colMeans(eval[odd_index, , drop = FALSE], na.rm = TRUE)
  g1 <- colMeans(eval[even_index, , drop = FALSE], na.rm = TRUE)
  
  tmp <- rbind(g0, g1) %>% as.data.frame()
  tmp$Group  <- rownames(tmp)
  tmp$Method <- method
  rownames(tmp) <- NULL
  tmp <- tmp[, c("Method", "Group", setdiff(names(tmp), c("Method", "Group")))]
  
  model_eval_list[[length(model_eval_list) + 1L]] <- tmp
}

model_eval <- dplyr::bind_rows(model_eval_list)
model_eval
```
